Trapdoors and Random Walks With LWE

### 2017-06-28 latex/trapdoors_randomwalks/paper.tex
renamed variables

\author{Ella Rose}
\institute{Paso Robles, CA \email{python_pride@protonmail.com}}
\title[LWE Trapdoor One-Way Function]{LWE Trapdoor One-Way Function}

@@

% use optional argument because the \LaTeX command breaks the PDF keywords
\keywords[Homomorphic Encryption, Public Key Cryptography]{Homomorphic Encryption, Public Key Cryptography}

\begin{abstract}
 We define a simple trapdoor function based on the Learning-With-Errors problem in one dimension, and present a key exchange algorithm based on the trapdoor. The trapdoor algorithm is efficient in terms of key size, key generation time, ciphertext size, and public/private key operation time: We can perform many thousands of public and private key operations per second with a naive python implementation on modest hardware. 
 \end{abstract}

@@

\section{Introduction}
A Trapdoor One-Way Function (TOWF) is a function that is easy for anyone to compute in one direction, but is hard to compute in reverse without some additional piece of information. A TOWF makes exchanging cryptographic keys a straightforward exercise: Suppose that Bob happens to know the trapdoor information for a TOWF: Alice can simply apply the TOWF to a randomly generated key and send the output to Bob. Bob can then use the trapdoor information to invert the output and obtain the key.

The Learning-With-Errors (LWE) problem is a mathematical problem that is believed to be difficult to solve efficiently even with a quantum computer.

@@

@@

\section{Algorithm}
There is a publicly agreed upon modulus $N$. The same $N$ may always be used. It should be prime; It it is not secret, and otherwise only needs to be both agreed upon and large enough. The size of $N$ is determined by the rest of the parameter sizes which are discussed below; In our python implementation which targets a 256-bit security level, $N$ is slightly less then 800 bits in size. The larger $N$ is, the longer the BigNum arithmetic operations take, and the larger ciphertexts become; Consequently, we opt for the smallest $N$ possible.

@@

\section{Security Analysis}
There are two main goals: First, given the public key, the private key must not be obtainable; Second, The output of the public key operation must not be invertible without the private key.

\subsection{Key Generation}
It would appear that obtaining the private key from the public key must be done by guessing one of $p, p_i, r$. If this were not the case, then whatever algorithm that can be used to recover any of $p, p_i, r$ from $p r \mod N$ could be used to recover plaintexts from ciphertexts of the form $m k \mod N$ for random $m, k$ where one of $m, k$ is smaller then the other (but still large enough to reduce modulo $N$). It seems unlikely that such an attack exists and is faster then guessing $p, p_i,  r$.

@@

\subsection{Public Key Operation}
The public key operation appears to be a 1-dimensional version of the learning-with-errors problem.\\

@@

\subsection{Side Channel Analysis}
Assuming that BigNum operations can be implemented in a manner that does not leak timing information, it appears that a naive implementation of the algorithm is resistant to timing side channel analysis, as there are no tables or branches in the algorithm.

\section{Metrics}
Using our public domain python implementation that is configured for a 256-bit security level, we can perform 10,000 public key operations in slightly under 1 second. It requires .3 seconds to perform 10,000 private key operations, which implies that roughly 30,000 private key operations per second can be performed on very modest hardware. Our tests were performed on an Intel Celeron N2830 2.16GHz with 2 cores.

@@

\subsection{Implementation Complexity}
The implementation complexity of the scheme is very low. Our public domain python implementation of the algorithm spans barely 100 lines, including the documentation and comments.\\

@@

>>>>

\author{Ella Rose}
\institute{Paso Robles, CA \email{python_pride@protonmail.com}}
\title[Trapdoors and Random Walks With LWE]{Trapdoors and Random Walks With LWE}

@@

% use optional argument because the \LaTeX command breaks the PDF keywords
\keywords[Homomorphic Encryption, Public Key Cryptography, LWE]{Homomorphic Encryption, Public Key Cryptography, LWE}

\begin{abstract}
 We define a trapdoor function and random-walk type function based on the Learning-With-Errors problem; Together, the algorithms constitute post-quantum alternatives to traditional RSA and DH . Our constructions are practical and efficient in terms of size and space: public and private keys are small, ciphertext expansion is relatively low, and public and private key operation times are fast.
 \end{abstract}

@@

\section{Introduction}
A Trapdoor One-Way Function (TOWF) is a function that is easy for anyone to compute in one direction, but is hard to compute in reverse without some additional piece of information. A TOWF makes exchanging cryptographic keys a straightforward exercise: Suppose that Bob happens to know the trapdoor information for a TOWF: Alice can simply apply the TOWF to a randomly generated key and send the output to Bob. Bob can then use the trapdoor information to invert the output and obtain the key. With the right padding, a TOWF may be used for public key encryption as well; The difference is that instead of a random value, Alice can choose the value that she sends to Bob.

A Random-Walk function is a function that facilitates key agreement: Suppose Alice was to walk in a particular but random pattern where she ends up at a destination $A$; Bob does the same, walking in his own particular but random pattern to end up at a destination $B$. Then, if Alice starts at $B$ and walks her random path again, she will end up at a point $C$; If Bob starts at $A$ and walks his random path again, he too will end up at point $C$. In this way, both parties arrive at the same final location, and neither learns the particular way that the other one arrived. Elliptic-Curve Diffie-Hellman is an examplea of a Random-Walk type function.

The Learning-With-Errors (LWE) problem is a mathematical problem that is believed to be difficult to solve efficiently even with a quantum computer. Problems of this type take the form $a s + e \mod q$, where $a, q$ are public knowledge and $s, e$ are kept secret.

@@

Additionally, we also define a Random-Walk type function using the observation of the fact that LWE incorporates scaling and translation; The combined effects of scaling and translation act as the random steps on the path taken by Alice and Bob.

@@

\section{Trapdoor Algorithm}
There is a publicly agreed upon modulus $N$. The same $N$ may always be used. It should be prime; It it is not secret, and otherwise only needs to be both agreed upon and large enough. The size of $N$ is determined by the rest of the parameter sizes which are discussed below; In our python implementation which targets a 256-bit security level, $N$ is slightly less then 800 bits in size. The larger $N$ is, the longer the BigNum arithmetic operations take, and the larger ciphertexts become; Consequently, we opt for the smallest $N$ possible.

@@

\subsection{Security Analysis}
There are two main goals: First, given the public key, the private key must not be obtainable; Second, The output of the public key operation must not be invertible without the private key.

\subsubsection{Key Generation}
It would appear that obtaining the private key from the public key must be done by guessing one of $p, p_i, r$. If this were not the case, then whatever algorithm that can be used to recover any of $p, p_i, r$ from $p r \mod N$ could be used to recover plaintexts from ciphertexts of the form $m k \mod N$ for random $m, k$ where one of $m, k$ is smaller then the other (but still large enough to reduce modulo $N$). It seems unlikely that such an attack exists and is faster then guessing $p, p_i,  r$.

@@

\subsubsection{Public Key Operation}
The public key operation appears to be a 1-dimensional version of the learning-with-errors problem.\\

@@

\subsubsection{Side Channel Analysis}
Assuming that BigNum operations can be implemented in a manner that does not leak timing information, it appears that a naive implementation of the algorithm is resistant to timing side channel analysis, as there are no tables or branches in the algorithm.

\section{Random-Walk Algorithm}
There is a publicly known and agreed upon starting point $x$, as well as another starting point and modulus $y$. The size of $y$ determines the ciphertext size and key size; The size of $x$ should be close to but smaller then $y$. We elect to utilize 256 bits for the size of $x$, and $264$ bits for the size of $y$; This ensures that 256 bits of key material are available as a shared secret.

\subsection{Basic Concept}
A random walk requires a starting point, as well as directions for where to proceed next. The LWE expression $a s + e \mod q$ incorporates a scaling by $s$ and a translation by $e$; These steps serve as the directions for the random walk, as they determine which destination point is selected; The value of $a$ acts as the initial location where the walk begins from.

\subsection{Random Walk Algorithm}
\begin{flalign*}
point_2 := point_1 s + e y \mod y
\end{flalign*}

The above expression calculates the location of the next point (which may be the destination) according to the values of $s, e$; The value of $e$ is mutliplied by $y$ to facilitate the reduction modulo $y$. This expression is used to generate a public key, as well as to compute the shared secret between parties.

\subsection{Key Generation}
A private key consists of random integers $s, e$. A public key consists of the point that is arrived at after using the private key "directions" $s, e$ to walk, starting from the base points $x, y$:

\begin{flalign*}
s := random(s_{size})\\
e := random(e_{size})\\
point := x s + e y \mod y\\
key_{private} := s, e\\
key_{public} := point
\end{flalign*}

\subsection{Key Agreement}
A shared secret can be obtained by applying the random walk function using the other parties public key as the starting location.

\begin{flalign*}
secret := key_{public} s + e y \mod y
\end{flalign*}

\subsection{Additional Points on the Walk}
In the examples given in the prior section, the random walk function is only applied once; In fact, it may be applied many times. An arbitrary number of points can be walked through without increasing space requirements - it only costs more time to walk further.

It may be required that a greater number of points on the path is required for security; We are unable to demonstrate an attack against a single point walk. We choose to utilize the fewest number of points required for efficiency reasons provided security can still be maintained.

\section{Metrics}
\subsection{Trapdoor Algorithm}
Using our public domain python implementation that is configured for a 256-bit security level, we can perform 10,000 public key operations in slightly under 1 second. It requires .3 seconds to perform 10,000 private key operations, which implies that roughly 30,000 private key operations per second can be performed on very modest hardware. Our tests were performed on an Intel Celeron N2830 2.16GHz with 2 cores.

@@

\subsubsection{Implementation Complexity}
The implementation complexity of the scheme is very low. Our public domain python implementation of the algorithm spans barely 100 lines, including the documentation and comments.\\

@@

\subsection{Random-Walk Algorithm}
